Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-07-19 decorrelation/decorrelation.tex
updating, including references, cleaning up a bit

Another assumption that is instrumental in proving many threshold theorems is Markovianity \cite{Kitaev1997}. There are many ways in which devices can behave in non-Markovian ways, and NISQ devices have already been shown to be subject to drift in controls, \cite{Kelly2018} and in general non-Markovian noise \cite{BlumeKohout2017}. Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to try to remove correlated noise. Indeed, other have even implemented these routines natively in hardware \cite{Ware2018}. These routines all require some level of sophistication. In the case of \cite{Campbell2017}, the analysis assumes Solovay-Kitaev [CITATION]-like decomposition of unitaries, and treats Pauli and Clifford gates as being free to implement, which is far from the reality of modern quantum computers. In \cite{Wallman2016} and the implementation in \cite{Ware2018}, the random frames are implemented in each sequence by precompiling the inserted Pauli frame rotations with the Clifford operations. In general, this cannot be done efficiently for non-Clifford operations, and in addition, this introduces a linear-time precompilation overhead for shot of the circuit.

@@

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome, [CITATION NEEDED], and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be massively reduced, and the process can be made much closer to Markovian, at no real cost to the fidelity of the implemented gate. Unlike \cite{Ware2018}, \cite{Wallman2016}, and \citep{Campbell2017}, this routine can be implemented in the firmware of the quantum computer, and requires only an instruction for a random choice - no pre or post processing is required.

@@

This can be solved with a constrained minimization algorithm, such as Sequential Least Squares Programming.[CITATION NEEDED]

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires the use of a convex solver, and in particular does not have the restriction that the optimal channel be decomposable into a given family of controls. Therefore we can view our approach as a regularization of this method, where we have imposed the condition that our final channel must arise from a BCS.[IS THIS ACTUALLY A REGULARIZATION? DOES IT IMPOSE SOME KIND OF SPARSITY] In doing so, we move from using a convex optimization routine to compute the diamond norm to needing to computing a sum. In the next section we give a simple example, followed by numerical results for one qubit and two qubits gates in sections \ref{1Q Gates} and \ref{2Q Gates}.

@@

In this example we generated 10 candidate pulses using the GRAPE algorithm with a high convergence tolerance, with the intention that the optimizer would find solutions with some degree of noticeable coherent error. For each pulseshape, which enacts some unitary $U_i$, we performed an RPE-like experiment [REF RPE] where the $\sigma_y$ and $\sigma_z$ expectations values were measured at the end of a sequence of $n$ applications of $U_i$, where n ranged from 1 to 100, and each circuit was sampled 1000 times. The results are plotted in [REF SUBPLOTS OF \ref{fig:rebit}]. Then, using the same optimization tool chain as in \ref{numerical}, we picked out the coeffients $p_i$ such that $p_i, U_i$ are a BCS, and performed the same experiment, but where for each shot of the experiment, we precompiled a different circuit. In particular, in our sampling procedure, circuit $U_{i_0}, U_{i_1}, ..., U_{i_{n-1}}$ occurs with probability $p_{i_0}p_{i_1}...p_{i_{n-1}}$. While in this example high fidelity, low coherent error controls can be produced [REFERENCE SUBPLOT OF \ref{fig:pulseshapes}], this demonstrates that this method can reduce coherent errors drastically.

@@

\section{Conclusion}
We have shown numerically and experimentally that drawing from a collection of implementations of a gate with the correct probabilties can reduce coherent error by an order of magnitude, while increasing incoherent error by a smaller amount, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing quantum hardware. Future directions for this work include moving the random gate selection from a precompilation step onto runtime FPGA logic, investigating other optimzation routines such as [GOAT, CRAB], and using more precise benchmarking routines such as GST [REF GST] to more quantitatively investigate the performance of these routines. The code used in this paper is available at [REF MY GITHUB].

%It might be interesting to show a dhistogram of the pribabilities assigned to each control to show that they aren't all zero. and maybe

>>>>

Another assumption that is instrumental in proving many threshold theorems is Markovianity \cite{Kitaev1997}. There are many ways in which devices can behave in non-Markovian ways, and NISQ devices have already been shown to be subject to drift in controls, \cite{Kelly2018} and in general non-Markovian noise \cite{BlumeKohout2017}. Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to try to remove correlated noise. Indeed, other have even implemented these routines natively in hardware \cite{Ware2018}. These routines all require some level of sophistication. In the case of \cite{Campbell2017}, the analysis assumes some algorithm for decomposing arbitray unitaries, and treats Pauli and Clifford gates as being free to implement in those decompositions, which is far from the reality of modern quantum computers. In \cite{Wallman2016} and the implementation in \cite{Ware2018}, the random frames are implemented in each sequence by precompiling the inserted Pauli frame rotations with the Clifford operations. In general, this cannot be done efficiently for non-Clifford operations, and in addition, this introduces a linear-time precompilation overhead for shot of the circuit.

@@

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be massively reduced, and the process can be made much closer to Markovian, at no real cost to the fidelity of the implemented gate. Unlike \cite{Ware2018}, \cite{Wallman2016}, and \citep{Campbell2017}, this routine can be implemented natively in the control electronics of a quantum computer, and only requires that the onboard sequencer has the abiltiy to generate random numbers - no pre or post processing is required.

@@

This can be solved with a constrained minimization algorithm, such as Sequential Least Squares Programming\cite{wright1999numerical}.

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires the use of a convex solver, and in particular does not have the restriction that the optimal channel be decomposable into a given family of controls. Our routine, on the other hand, uses a convex optimization routine to compute the diamond norm to needing to computing a sum. In the next section we give a simple example, followed by numerical results for one qubit and two qubits gates in sections \ref{1Q Gates} and \ref{2Q Gates}.

@@

In this example we generated 10 candidate pulses using the GRAPE algorithm with a high convergence tolerance, with the intention that the optimizer would find solutions with some degree of noticeable coherent error. For each pulseshape, which enacts some unitary $U_i$, we performed an RPE-like experiment\cite{Kimmel2015} where the $\sigma_y$ and $\sigma_z$ expectations values were measured at the end of a sequence of $n$ applications of $U_i$, where n ranged from 1 to 100, and each circuit was sampled 1000 times. The results are plotted in [REF SUBPLOTS OF \ref{fig:rebit}]. Then, using the same optimization tool chain as in \ref{numerical}, we picked out the coeffients $p_i$ such that $p_i, U_i$ are a BCS, and performed the same experiment, but where for each shot of the experiment, we precompiled a different circuit. In particular, in our sampling procedure, circuit $U_{i_0}, U_{i_1}, ..., U_{i_{n-1}}$ occurs with probability $p_{i_0}p_{i_1}...p_{i_{n-1}}$. While in this example high fidelity, low coherent error controls can be produced [REFERENCE SUBPLOT OF \ref{fig:pulseshapes}], this demonstrates that this method can reduce coherent errors drastically.

@@

\section{Conclusion}
We have shown numerically and experimentally that drawing from a collection of implementations of a gate with the correct probabilties can reduce coherent error by an order of magnitude, while increasing incoherent error by a smaller amount, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing quantum hardware. Future directions for this work include moving the random gate selection from a precompilation step onto runtime FPGA logic, investigating other optimzation routines such as CRAB \cite{Caneva2011} and GOAT\cite{Machnes2018}, and using more precise benchmarking routines such as GST\cite{BlumeKohout2017} to more quantitatively investigate the performance of these routines. The code used in this paper is available at \cite{decorrelating_errors}.

\section{Acknowledgements}
%It might be interesting to show a dhistogram of the pribabilities assigned to each control to show that they aren't all zero. and maybe
